---
title: "Towards open collaboration in biomedical and health research: An open collaborative scoping review"
date: last-modified
author:
  - name: Mario Garcia
    affiliations:
      - ref: ku
  - name: Daniel B. Ibsen
    affiliations:
      - ref: sdca
      - ref: au
  - name: Luke W. Johnston
    affiliations:
      - ref: sdca
      - ref: au
affiliations:
  - id: sdca
    name: Steno Diabetes Center Aarhus
    country: Denmark
  - id: au
    name: Aarhus University
    country: Denmark
  - id: ku
    name: Copenhagen University
    country: Denmark
format:
  html: default
  academic-typst:
    section-numbering: "1.1.1"
    columns: 2
    papersize: a4
    margin:
      x: 2cm
      y: 2cm
format-links:
  - html
  - format: academic-typst
    text: PDF
    icon: file-pdf
abstract: |
  **Background**: Greater, more extensive collaboration is increasingly needed
  in the modern research environment, often to the level of team-based science.
  Combined with this need is the emphasize on doing more open science. However,
  there seems to be less attention on the overlap between open science and
  collaboration.

  **Objectives**: To conduct a scoping review on open collaborative practices used 
  across academia, including methods, tools, and resources.

  **Eligibility criteria**: We included any document that discussed, in any way,
  open collaboration practices, only those in English, and only those that were
  open access. We searched 5 years from whenever we ran a search.

  **Sources of evidence**: We used the OpenAlex database to search for scholarly
  works, which includes a large number of sources. Our search period was
  between 2019-02-28 to 2024-02-29.

  **Charting methods**: All three authors participated in the search, extraction,
  and charting process. As no standard charting form exists for our purposes,
  we developed our own template.

  **Results**: TBD

  **Conclusions**: TBD
execute: 
  echo: false
  message: false
  warning: false
---

```{r setup}
#| include: false
knitr::opts_chunk$set(echo = FALSE)
library(targets)
library(readr)
library(dplyr)
library(glue)
library(stringr)
library(here)
source(here("R/openalex-search.R"))
search_terms_inline <- glue("({search_terms()$title_search}[title]) AND {search_terms()$general_search}")
reviewed_records <- tar_read(reviewed_article_records)

list_refs <- bib2df::bib2df(here("doc/reviewed-records.bib")) |>
  filter(CATEGORY != "COMMENT") |>
  glue_data("@{BIBTEXKEY}") |>
  str_c(collapse = ";") |>
  append("[", 0) |>
  append("]") |>
  str_c(collapse = "")
```

## Background

The time when a single scientist could work alone in a lab and make
meaningful progress in their field is long gone. Scientific research in
the modern era, especially in certain fields, requires working with
other people, oftentimes a large number of people [@Forscher2022].
Research questions are increasingly more challenging to answer as they
require more time, resources, expertise, and personnel in order to
meaningfully answer them.

With the current needs and limitations when doing scientific research,
collaborations range from those within a single group or lab, to a few
researchers across different departments in the same institution, to
many researchers and personnel spread across multiple research centers.
Collaborations are needed to address basic limitations of working with
small or constrained populations, to sharing valuable but expensive
equipment, or to assisting with technical expertise or domain knowledge.

At the same time, there is also a higher need for science to be more
open, higher quality, and more rigorous [@Cole2024; @Umbach2024;
@Korbmacher2023], which necessitates that researchers collaborate with
more researchers and with a more diverse group of researchers to fulfill
these needs. Open scientific practices, such as open data, open source,
and open materials, aim to improve the reliability and accessibility of
scientific output to increase the societal impact of the results.
There's greater attention and scrutiny towards verification of results,
a core tenet of science, through replication and reproducibility. In
many ways, the "replication crisis" in science has been a driving force
behind the push for more open science and "team science" practices that
necessitate more collaboration [@Korbmacher2023].

There are many claimed benefits to openness in science, including in
collaboration [@Forscher2022; @Kavanagh2021]. For instance:

-   Greater transparency and accountability, because the built-in
    openness makes it easy to see what was done, how, and why.
-   Better reproducibility and inspectability, because everything can be
    easily viewed and can be verified, which is a key tenet of science.

The intersection between greater collaboration and more openness
highlights key challenges researchers face. How can we effectively
collaborate together in a way that fulfills the demand for more
openness, transparency, and accountability, while at the same time also
not substantially impacting our productivity (or ideally improving it).
How can we collaborate in an open and transparent way? What are the best
practices and tools we can use? What is an ideal collaborative workflow
and how close or how far are we from this ideal in reality?

From our own personal experience working in the field of health
research, we have varied and diverse ways of collaborating and of being
open. Often the most commonly used approach to collaboration is emailing
Word documents around to get feedback and using meetings to discuss and
agree on things. Rarely do we systematically examine whether we are
following best practices for collaboration nor do we collaborate in an
open way following the principles of open science. As for open science,
while it was first introduced in 1985 [@Chubin1985] and discussed in
academic outputs regularly since then, the actual practices of being
open in science have been minimal or largely confined to specific
sub-branches of openness like open access.

This lack of visible role-models or standards of practice for
effectively collaborating and actively working openly presents a clear
barrier to uptake of these practices. This has real consequences to
continued productivity and quality of scientific research, which is
increasingly apparent in the modern era. These consequences are even
more apparent when comparing to more technically-inclined communities
such as those in open source software. These communities have massively
benefitted from integrating more openness into their workflows, to the
extent that the vast majority of the world's digital infrastructure is
built on and runs open source software. This highlights a greater need
for a paradigm shift in how we researchers collaborate to be able to
effectively and efficiently tackle the challenges of the modern era.

Open collaboration has many potential benefits, of which most overlap
with the benefits of open science. For instance, open collaboration can
lead to:

-   Easier onboarding of new collaborators, since everything is open and
    easily findable.
-   Higher collaboration opportunities, because more openness means more
    visibility as more parts of the scientific workflow are
    disseminated.
-   Faster time to dissemination, because the built-in openness makes it
    easier to disseminate results.

Unfortunately, there are, to our knowledge, very few resources and
examples of how to integrate open science into collaborations and how to
make collaboration be open [@Korbmacher2023].

### Objectives

Our aim is to conduct a scoping review to provide an overview of current
practices of open collaboration in the academic world. For this scoping
review, we define open collaboration using the definition as found in
[@Forte2013]:

> "*an online environment that (a) supports the collective production of
> an artifact (b) through a technologically mediated collaboration
> platform (c) that presents a low barrier to entry and exit and (d)
> supports the emergence of persistent but malleable social
> structures.*"

The specific aims of this scoping review are to:

1.  Provide an overview of current practices of or opinions about
    research collaboration that follow basic open principles (e.g.,
    transparency, accessibility)

2.  Summarize existing online tools and resources available to improve
    open collaboration in research

We've expanded on our original aims to include a **secondary aim** of
building an open source R-based pipeline for conducting scoping reviews.
The entire source code, as well as text and collaboration workflows, are
found on our GitHub repository
[`science-collective/scoping-review`](https://github.com/science-collective/scoping-review).

## Methods

The full protocol for this scoping review was uploaded to the Open
Science Framework [@Johnston2022]. As with any research project, things
evolve compared to what was originally intended or planned. In this
section we briefly re-state what we described in the protocol and
especially we describe what was changed from the protocol.

This is a scoping review, so we followed the framework described in
[@Arksey2005; @Levac2010] as well as the guidelines outlined in the
PRISMA-ScR statement [@Tricco2018].

### Deviations and challenges

We originally aimed to review individual archives that included PubMed,
bioRxiv, Scopus, and several others (the full list is in the protocol),
and to use R packages with web API connections to each of these sources
to programmatically extract the sources we wanted. However we
encountered substantial issues that completely changed how we actually
found and extracted the sources.

The first challenge we encountered was that, while most of the source
databases had (hypothetically) exposed APIs along with R packages
available to access them, they didn't always work well or had
complicated instructions for actually using them. For instance, the
preprint repositories bioRxiv and medRxiv didn't have an up to date R
package to access the preprints nor was the web API well described, nor
well designed. In order to effectively use it, you need to download the
entire database of preprints locally before being able to search for the
preprints you want. This made it effectively impossible to use.

The second challenge we encountered was a difference in results between
using the web API compared to using the web search interface for some
source databases. There were some small differences in search results
between the API query compared to the web-based query. We couldn't
identify or understand how these differences happened or way, though
they were quite small differences (less than 10 out of thousands of
scholarly works from the search results).

The third challenge we had was that some source databases, like the Web
of Science or Scopus, didn't work consistently between the authors. In
order to use the web API with the R packages, you need to generate a
token for authentication. However, for some authors, the token worked
fine, and for others, it didn't work at all, with no explanation or
error message. Given our interest in and aim for reproducibility, since
the same code ran differently between authors, we decided to exclude
these source databases.

We also originally intended on searching websites, blogs, and other
online resources, but it was very difficult to programmatically and
systemtically achieve this. So we ended up only searching for scholarly
work that were indexed in databases.

Given these challenges however, we unintentionally found
[OpenAlex](https://openalex.org/), which is a public database of
scholarly output that aggregates dozens of other source databases into
one, easy interface. This resource also has an R package called
`{openalexR}` [@Aria2024] with a very well designed interface, which
simplified or made redundant much of our code we wrote. It also made it
unnecessary to use the individual R packages for each source database
that we originally listed in the protocol. By using this archive, it
resolved all our challenges and barriers we had before. For a full list
of where they get their source data from, see their ["About the
data"](https://help.openalex.org/hc/en-us/articles/24397285563671-About-the-data)
page.

### Document selection

We developed the initial search strategy in consultation with a research
librarian. We collected the data via systematic searches of databases.
All authors were involved in reviewing the collected sources, from
reviewing the titles, to abstracts, and finally to full-text documents.

#### Information sources

We used OpenAlex, after encountering the challenges described above,
which is an open database with scholarly works from Microsoft Academic
Graph, Crossref, ORCID, ROR, DOAJ, Unpaywall, Pubmed, Pubmed Central,
the ISSN International Centre, Internet Archive, Web crawls,
subject-area and institutional repositories from arXiv and Zenodo.

#### Search terms

We used the following search terms when searching for scholarly work:

> `r search_terms_inline`

#### Inclusion and exclusion criteria

Inclusion criteria included any document where open collaboration
practices are not the primary focus. We relied on the definition of open
collaboration from [@Forte2013] in determining whether the records were
relevant. Any published document with reporting on current open
collaboration practices. Any published document with advice, guidance,
tools, and/or recommendations for improving open collaboration. Article
language in English. Exclusion criteria were documents that do not
report on specific open collaboration practices.

While we didn't change our search terms for the initial search and
extraction, we found that our search terms were not precise enough and
got many irrelevant scholarly works. The full R code to filter down the
search results is kept in the
[`R/exclusions.R`](https://github.com/science-collective/scoping-review/blob/main/R/exclusions.R)
file on the GitHub repository. For example, we extracted a large number
of scholarly works describing surgery such as "open wound", electronics
such as "open circuit", or environmental such as "open water" that we
had to post-process and exclude.

#### Charting procedure

At least two of us extracted data using a standardized and tested
template. Data regarding the data source (e.g., author, title,
publication year), open collaboration practices, and any other relevant
information, will be extracted. Extracted data will be summarized with
the descriptive analytical method described in [@Arksey2005], which is
aimed at identifying and summarizing different open collaboration
practices.

#### Search period

We were not able to dedicate as much time to this project as we had
initially planned, so our "stopping rule" for end date didn't match with
our actual search period. The actual end date for when we did the last
search results extraction was
[2024-02-29](https://github.com/science-collective/scoping-review/commit/f288d3f23d028e77ad1081e0cfe3583ce4dcb8b6).

The start date remained 5 years from our end date, which would have been
approximately 2019-02-28.

## Results

```{r}
#| echo: false
n_papers <- tar_read(records_title_only) |>
  nrow()

n_papers_title_review <- tar_read(records_after_title_exclusion) |>
  nrow()

n_papers_excluded_title_review <- n_papers - n_papers_title_review

n_titles_selected <- tar_read(titles_selected) |>
  nrow()

n_paper_full_text <- tar_read(abstracts_selected) |>
  nrow()

n_papers_selected_for_review <- read_csv(here("data", "article_characteristics.csv")) |>
  nrow()

# get table of study characteristics
article_characteristics <- read_csv(here("data", "article_characteristics.csv"))

n_guides <- article_characteristics |>
  filter(paper_type == "guide") |>
  nrow()

n_examples <- article_characteristics |>
  filter(paper_type == "example") |>
  nrow()

n_surveys <- article_characteristics |>
  filter(paper_type == "survey") |>
  nrow()

n_org <- article_characteristics |>
  filter(str_detect(themes, "organization")) |>
  nrow()

n_tools <- article_characteristics |>
  filter(str_detect(themes, "tools")) |>
  nrow()

n_education <- article_characteristics |>
  filter(str_detect(themes, "education")) |>
  nrow()

n_researcher <- article_characteristics |>
  filter(relevant_for == "researcher") |>
  nrow()

n_research_group <- article_characteristics |>
  filter(relevant_for == "research group") |>
  nrow()

n_research_center <- article_characteristics |>
  filter(relevant_for == "research center") |>
  nrow()

n_research_centers <- article_characteristics |>
  filter(relevant_for == "multiple centers") |>
  nrow()

n_research_field <- article_characteristics |>
  filter(relevant_for == "research field") |>
  nrow()
```

Our search yielded a total of `r n_papers` papers from which we filtered
`r n_papers_excluded_title_review` based on key words. Three reviewers
independently screened for the remaining `r n_papers_title_review`
titles from which we screened `r n_titles_selected` abstracts. Of these,
we assessed `r n_paper_full_text` full-text papers and we included
`r n_papers_selected_for_review` in the scoping review. These were
`r list_refs`. Two of the documents, while they passed our abstract
review, did not contain anything relevant when we reviewed the full text
[@Network2023a; @Yordanov2022], so we did not continue with charting
them.

The identified papers included `r n_guides` guides, `r n_examples`
examples and `r n_surveys` surveys.

`r n_org` of the `r n_papers_selected_for_review` papers included
recommendations for organizational practices to improve open
collaboration, `r n_tools` mentioned specific tools used for open
collaboration and `r n_education` highlighted educational needs. The
identified papers discussed open collaboration on different
organizational levels. `r n_researcher` were mostly directed at the
individual researcher, `r n_research_group` at a research group,
`r n_research_center` at a research center, `r n_research_centers` at
several research centers and `r n_research_field` at entire research
fields.

### Organizational practices of open collaboration

A common theme across organizational levels was the need for
infrastructure that promotes open collaboration, building a community of
collaborators and using online tools in accordance with the FAIR
(Findable, Accessible, Interoperable and Reusable) principles.

individual level - add text

laboratory/research group level - add text

For center-level one paper described how they transitioned their
neuroimaging research center toward more open and collaborative
practices. Their central focus was on creating more open code and data
to improve open collaboration [@Bush2022]. First, they used open-source
code, then standardized data formats, developed analysis pipelines,
shared data openly and developed protocols for data documentation. While
the organizational transition to improve open collaboration was
described, little was mentioned about how this has improved
collaborative work at the research center.

Interestingly, the level of organization with most papers was the
multi-center level with two papers from ManyPrimates [@Primates2019;
@ManyPrimates2021] and the Reproducibility Network [@Network2023]. In
this organization level the most urgent needs are: 1) identifying the
needs for open collaboration and 2) creating an infrastructure that
allows those needs to be satisfied. Indeed, Grange et al describes a
systematic review as an exploration of the needs for 11 UKRN centers to
implement open collaborative practives. In this paper they specially
focus on tools and resources among the 11 institutions and make clear
the need for starting to utilize open resources as a first step to adapt
open collaborative practices. The first paper from ManyPrimates
[@Primates2019] describes the start of another multi-center
collaboration, albeit from different institutions. It is interesting to
observe that in this situation, the complex issue is to start the
collaboration itself, which was ignited thanks to a symposium and an
email chain among future collaborators. In contrast, ManyPrimates could
quickly develop an open collaborative structure with relative ease in
comparison to changing an already solid infrastructure, like in the case
of @Network2023.

research field level

Here, we provide the specific needs for implementing open collaboration
in each organizational level, as well as describing the differences in
approaches among experiences between and across levels (Figure 2).

### Useful tools for open collaboration

### Educational needs

## Discussion

### Conclusions

## Contributions

<!-- TODO: add contributions according to: https://credit.niso.org/ -->

We originally had another author involved, HC, who is on the protocol
author list. Since the time we uploaded the protocol, she has moved out
of academia and could no longer participate in any work.

Otherwise, the remaining authors (MG, DBI, and LWJ) had the following
contribution roles, following the [CRediT](https://credit.niso.org/)
taxonomy of conceptualization, data curation, formal analysis,
investigation, methodology, writing - original draft, writing -
reviewing and editing, project administration, software, validation, and
visualization.
